Simulations (mix normal prior), ctwas_rss, 45k samples

### 2021-02-19 analysis/index.Rmd
mix normal simulate phenotype

@@

>>>>

### Using normal distribution as prior distribution for causal effects

@@

### Using normal mixture distribution as prior distribution for causal effects

* [Using UKbiobank data (22k samples), fusion weights Adipose](simulation-ctwas-ukbWG-gtex.adipose_s40.22_mixnormal.html)

* [Using UKbiobank data (45k samples), fusion weights Adipose](simulation-ctwas-ukbWG-gtex.adipose_s80.45_mixnormal.html)

### 2021-02-19 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22_mixnormal.Rmd
mix normal simulate phenotype

---
title: "Simulations to test ctwas summary stats version, 22k samples"
output: html_document
---

@@

```{r files}
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22.2_pgenfs.txt"
outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210113/" # /
runtag = "ukb-s40.22-adi"
configtags = 1
simutags = paste(rep(1:2, each = length(1:5)), 1:5, sep = "-")

@@

      To simulate phenotype data, first we impute gene expression based on expression models, then we set gene/SNP pi1 and PVE, get rough effect size for causal SNPs and genes and simulate phenotype under the sparse model with spike and slab prior.
      Then we performed GWAS for all SNPs and get z scores for each by univariate linear regression.

@@

2. Run ctwas_rss
We used LDetect to define regions. `ctwas_rss` algorithm first runs on all regions to get rough estimate for gene and SNP prior. Then run on small regions (having small probablities of having > 1 causal signals based on rough estimates) to get more accurate estimate. To lower computational burden, we downsampled SNPs (0.1), estimate parameters and convert back to orginal scale. Lastly, run susie with given L for all regions and for all genes and SNPs using estimated prior and prior variance.

@@

## Results-last step impact

After we have the estimated parameters, we obtain PIP for genes in one the following ways:
(1) run susie for each region using all SNPs
```{r param1-1, fig.height=3, fig.width=8, message=F, cache=T}
configtag <- 1
runtag2 = "config1_ctwas_rss_s3_full/ukb-s40.22-adi"

simutags <- paste(1, c(2,4,5), sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag2, simutags)

simutags <- paste(2, c(2,4), sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag2, simutags)
```
(2) run susie for each region thinned SNPs 
```{r param1-2, fig.height=3, fig.width=8, message=F}
configtag <- 1
runtag2 = "config1_ctwas_rss_s3_thin/ukb-s40.22-adi"

simutags <- paste(1, c(2,4,5), sep = "-")
plot_PIP(configtag, runtag2, simutags)

simutags <- paste(2, c(2,4), sep = "-")
plot_PIP(configtag, runtag2, simutags)
```

(3) run susie for each regions with thinned SNPs, then select regions with strong gene PIP (max PIP > 0.8) and rerun susie with full SNPs.
```{r param1-3, fig.height=3, fig.width=8, message=F}
configtag <- 1
runtag2 = "ukb-s40.22-adi"

simutags <- paste(1, c(2,4,5), sep = "-")
plot_PIP(configtag, runtag, simutags)

simutags <- paste(2, c(2,4), sep = "-")
plot_PIP(configtag, runtag2, simutags)
```

## Results-more simulation settings
We run the last step of `ctwas_rss` using way (3) as described above.

```{r param2, fig.height=3, fig.width=8, message=F}

@@

@@

simutags <- paste(9, 1:4, sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)

>>>>

---
title: "Simulations (mix normal prior), ctwas_rss, 22k samples"
output: html_document
---

@@

```{r files}
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22.2_pgenfs.txt"
outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210215/" # /
runtag = "ukb-s40.22-adi"
configtags = 1
simutags = paste(rep(1:2, each = length(1:5)), 1:5, sep = "-")

@@

      To simulate phenotype data, first we impute gene expression based on expression models, then we set gene/SNP pi1 and PVE, get rough effect size for causal SNPs and genes and simulate phenotype under the mixture of normal distributions.

* The prior distribution for causal effect: 
```{r }

nmix <- 4
pi_k <- rep(1/nmix, nmix)

ra <- 2**(1: length(pi_k))
sa2_kr <-  ra / (pi_k * sum(ra))

cat("number of mixtures: ", nmix, "\n")
cat("variance of each component:", sa2_kr, "\n")

hist(mixtools::rnormmix(2000, lambda = pi_k,
                        mu = rep(0, length(pi_k)),
                        sigma = sa2_kr), breaks = 100, main = "prior distribution", xlab = "")
```

Then we performed GWAS for all SNPs and get z scores for each by univariate linear regression.

@@

2. Run ctwas_rss
We used LDetect to define regions. `ctwas_rss` algorithm first runs on all regions to get rough estimate for gene and SNP prior. Then run on small regions (having small probablities of having > 1 causal signals based on rough estimates) to get more accurate estimate. To lower computational burden, we downsampled SNPs (0.1), estimate parameters. We run susie for all regions using estimated parameters, select regions with strong gene signals (max gene PIP > 0.8) and rerun using full SNPs for these regions.

@@

```{r param, fig.height=3, fig.width=8, message=F}

@@

simutags <- paste(1, 1:5, sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)

simutags <- paste(2, 1:5, sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)

@@

simutags <- paste(9, 1:5, sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)

### 2021-02-19 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_mixnormal.Rmd
mix normal simulate phenotype

---
title: "Simulations (mix normal prior), ctwas_rss, 22k samples"
output: html_document
---

@@

```{r files}
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22.2_pgenfs.txt"
outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210215/" # /
runtag = "ukb-s40.22-adi"
configtags = 1
simutags = paste(rep(1:2, each = length(1:5)), 1:5, sep = "-")

@@

```{r gene}
# p2 <- pow(niter, n, phenores[["batch"]][[1]][["sigma_beta"]], gene.p)
print(p2)
# save(p1,p2, file = "data/power_s40.22.Rd")
```

@@

plot_PIP <- function(configtag, runtag,  simutags){
   phenofs <- paste0(outputdir, "ukb-s40.22-adi", "_simu", simutags, "-pheno.Rd")
   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")

@@

```{r param, fig.height=3, fig.width=8, message=F}

configtag <- 1
runtag = "ukb-s40.22-adi"

>>>>

---
title: "Simulations (mix normal prior), ctwas_rss, 45k samples"
output: html_document
---

@@

```{r files}
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210216/" # /
runtag = "ukb-s80.45-adi"
configtags = 1
simutags = paste(rep(1:2, each = length(1:5)), 1:5, sep = "-")

@@

```{r gene}
# p2 <- pow(niter, n, phenores[["batch"]][[1]][["sigma_beta"]], gene.p)
print(p2)
# save(p1,p2, file = "data/power_s80.45.Rd")
```

@@

plot_PIP <- function(configtag, runtag,  simutags){
   phenofs <- paste0(outputdir, "ukb-s80.45-adi", "_simu", simutags, "-pheno.Rd")
   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")

@@

```{r param2, fig.height=3, fig.width=8, message=F}

configtag <- 1
runtag = "ukb-s80.45-adi"
